---
title: "Using temperature to describe stress"
author: "Mikaela M. Provost"
date: "7/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goals

Overal goal is to develop a method that defines stress through time based on a time series of temperature. Before running our stress model, I want to calibrate SST to match *in situ* temperature at depth. We have lots of *in situ* data, but we have so much more SST.

What data do we have?

* *In situ* daily temperature observations ( 9/12/2015 -- 8/17/2018) in two locations, Morro Prieto and Punta Prieta.

* Daily SST for a longer period of time (2009-2019) in both locations.

Update: randomly select one temperature per day from the 10 min data. Then run slope model on this time series of bottom temperature (instead of using mean daily T).



```{r error=FALSE, include=FALSE, message=FALSE, warning=FALSE}
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(grid)
```

```{r error=FALSE, include=FALSE, message=FALSE, warning=FALSE}
# --- Load data ---

# Try different data time series

# 1) April - Sept 2018 at Punta Prieta site
# d <- read.csv('C:/Users/Mikaela/Documents/GitHub/natividad/data/2018_04_15_Mexico_IslaNatividad_PuntaPrietaBuckets_1MAB_MiniDOT_SN095037.csv',header = TRUE, stringsAsFactors = FALSE)
# # create separate 'day' index column
# d$datetimeNew <- as.POSIXct(d$Date_Time_UTC,format="%m/%d/%Y %H:%M")
# d <- d %>% separate(Date_Time_UTC,into=c('dateUTC','timeUTC'),sep=' ') 
# d$dateUTC <- as.Date.factor(d[order(as.Date(d$dateUTC,format="%m/%d/%Y")),]$dateUTC,format="%m/%d/%Y")
# d$day <- as.numeric(as.factor(d$dateUTC))
# d <- d %>% select(dateUTC,Temperature_C,day)
# colnames(d) <- c("dateUTC","temp","dayindex")
# d$datatype <- rep("summer2018PP10min_insitu",length=length(d[,1]))
# head(d)
# 
# # reduce 10 min T data to mean daily T 
# davg <- d %>% group_by(dateUTC) %>% summarise(daymean=mean(temp))
# davg$dayindex <- seq(from=1,to=length(dayTmean$dateUTC),by=1)
# davg$datatype <- rep("summer2018PPdailyavg_insitu",length=length(davg[,1]))
# colnames(davg) <- c("dateUTC","temp","datatype")
# head(davg)
```


```{r error=FALSE, include=FALSE, message=FALSE, warning=FALSE}

# Date data frame: all possible days from 1/1/2003 - 12/27/2019
# use this to know if there are missing days, need it for averaging 
# observations on the same julian day at one site
datedf <- data.frame(date=seq(as.Date("2003-01-01"),as.Date("2019-12-27"),by="days"))
datedf$year <- format(as.Date(datedf$date, format="%d/%m/%Y"),"%Y")
datedf$dayindex <- rep(NA,length=length(datedf[,1]))
for(y in 1:length(unique(datedf$year))){ #for each yr
  yr = unique(datedf$year)[y] #add julian day in dayindex
  yrlgth = length(datedf[datedf$year==yr,]$year)
  datedf[datedf$year == yr,]$dayindex <- seq(from=1,to=yrlgth,by=1)}
head(datedf)
rm(yr,yrlgth,y)


# 2) 2009 - 2019 (Punta Prieta SST) 
# Brock filled in satelite SST derived data, missing bc cloud cover
d <- read.csv('C:/Users/Mikaela/Documents/GitHub/natividad/data/puntaprieta_temp.csv',header = TRUE, stringsAsFactors = FALSE)
d$datetime <- as.Date(d$datetime, "%m/%d/%Y")
names(d)[names(d)=="datetime"] <- "date"
d <- merge(d,datedf,by.x='date',by.y='date',all.x=T,all.y=T)
d$datatype <- rep("PP_SST",length=length(d[,1]))
d$flag <- as.factor(d$flag)
PP_SST <- d

# 3) 2003 - 2019 (Morro Prieto SST)
# Brock filled in satelite SST derived data, missing bc cloud cover
d <- read.csv('C:/Users/Mikaela/Documents/GitHub/natividad/data/morroprieto_temp.csv',header = TRUE, stringsAsFactors = FALSE)
d$datetime <- as.Date(d$datetime, "%m/%d/%Y")
names(d)[names(d)=="datetime"] <- "date"
d <- merge(d,datedf,by.x='date',by.y='date',all.x=T,all.y=T)
d$datatype <- rep("MP_SST",length=length(d[,1]))
d$flag <- as.factor(d$flag)
SST <- rbind(PP_SST,d)
rm(PP_SST,d)

# 4) 2015 - 2018 MP and PP in situ data 
d <- read.csv("C:/Users/Mikaela/Documents/GitHub/natividad/data/001_IslaNatividad_MorroPrietoPuntaPrieta_alldata_trimmed.csv",header=TRUE,stringsAsFactors = FALSE)
d$Date <- as.Date(d$Date,format="%m/%d/%Y")
names(d)[names(d)=="Date"] <- "date"
d <- d %>% filter(Site %in% c('MorroPrieto','PuntaPrieta'), Instrument_Type == 'MiniDOT')
d <- left_join(d,datedf,by="date")
in_situ <- d
rm(d)

# export some data for Will on my question:
# d <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
# #d$datatype <- rep("bottomT",length=length(d[,1]))
# names(d)[names(d)=="dailyT"] <- "bottomT"
# s <- SST %>% filter(datatype=="MP_SST") %>% select(date,temp)
# names(s)[names(s)=="temp"] <- "SST"
# head(s)
# ds <- left_join(d,s,by="date")
# ggplot(data=ds,aes(x=date,y=bottomT)) +
#          geom_line(color="blue") +
#   geom_line(aes(x=date,y=SST),color="black") +
#   ggtitle("SST (black) and bottom temp (blue)") +
#   ylab("T deg C")
# lines(ds$date,ds$SST)
# write.csv(ds,file="C:/Users/Mikaela/Documents/GitHub/natividad/data/forWill.csv")

```

#### Skip

 Calculate the average difference between SST and *in situ* temp

1. Find the average temperature per Julian day (at each site, for *in situ* and for SST). 

2. Calculate a smoothed monthly moving average. 

3. For each Julian day, calculate the difference between the moving average of SST and *in situ*. 

4. Use this daily-offset to adjust daily SST temperatures. 


```{r error=FALSE, include=FALSE, message=FALSE, warning=FALSE}
# generate time series of moving average 
window <- 30 #must be even number

# Morro Prieto in situ
# this line averages 10 min within each day
d <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(dayindex) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d$mm <- rep(NA,length(d$dayindex))
ts <- length(d$mm) - ((window/2)-1) - ((window/2)-1) -1 #start and end moving avg not at ends of time series
for(t in 1:ts){ #step through days
  d[((window/2)+t),]$mm <- mean(d$dailyT[t:(t+window-1)])} #average T over window}
d <- d[!is.na(d$mm),]
d <- d %>% gather("datatype","temp",2:3)
d$site <- rep("MP",length=length(d$temp))
MPmm <- d
rm(ts,t,d)
# Punta Prieta in situ
d <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(dayindex) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d$mm <- rep(NA,length(d$dayindex))
ts <- length(d$mm) - ((window/2)-1) - ((window/2)-1) -1 #start and end moving avg not at ends of time series
for(t in 1:ts){ #step through days
  d[((window/2)+t),]$mm <- mean(d$dailyT[t:(t+window-1)])} #average T over window}
d <- d[!is.na(d$mm),]
d <- d %>% gather("datatype","temp",2:3)
d$site <- rep("PP",length=length(d$temp))
# format julian day averages for in situ data
mmsit <- rbind(MPmm, d)
mmsit$datasource <- rep("insitu",length=length(mmsit$dayindex))
rm(ts,t,d,MPmm)
# plot site - julian day averages (can i add raw data??)
# ggplot(data=mmsit,aes(x=dayindex,y=temp,color=site))+
#   geom_point() + ggtitle("in situ: PP & MP monthly avg")

# Morro Prieto SST
d <- SST[SST$datatype == "MP_SST",] %>% group_by(dayindex) %>% summarize(dailyT = mean(temp,na.rm=TRUE))
d$mm <- rep(NA,length(d$dayindex))
ts <- length(d$mm) - ((window/2)-1) - ((window/2)-1) -1 #start and end moving avg not at ends of time series
for(t in 1:ts){ #step through days
  d[((window/2)+t),]$mm <- mean(d$dailyT[t:(t+window-1)])} #average T over window}
d <- d[!is.na(d$mm),]
d <- d %>% gather("datatype","temp",2:3)
d$site <- rep("MP",length=length(d$temp))
MPmm <- d

# Punta Prieta SST
d <- SST[SST$datatype == "PP_SST",] %>% group_by(dayindex) %>% summarize(dailyT = mean(temp,na.rm=TRUE))
d$mm <- rep(NA,length(d$dayindex))
ts <- length(d$mm) - ((window/2)-1) - ((window/2)-1) -1 #start and end moving avg not at ends of time series
for(t in 1:ts){ #step through days
  d[((window/2)+t),]$mm <- mean(d$dailyT[t:(t+window-1)])} #average T over window
d <- d[!is.na(d$mm),]
d <- d %>% gather("datatype","temp",2:3)
d$site <- rep("PP",length=length(d$temp))
PPmm <- d

# format julian day averages for SST data at both sites
mmsst <- bind_rows(MPmm,PPmm)
mmsst$datasource <- rep("SST",length=length(mmsst$temp))
rm(d,t,PPmm,MPmm,ts)

# julian day averages at both sites for SST & in situ 
mmjul <- bind_rows(mmsit,mmsst)
mmjul$sitesr <- paste(mmjul$site,mmjul$datasource,sep="_")
rm(mmsit,mmsst)

```



```{r error=FALSE, include=FALSE, message=FALSE, warning=FALSE}
# SST offset to match in situ bottom temperature

#Using the moving averages calculated above, find the difference between SST and in situ on each julian day.

# calculate: the average between the moving means of each data type 
# (in situ, SST) within each site (MP, PP)
# 1st: in situ data
d1 <- mmjul %>% filter(datatype=='mm' & datasource=='insitu' & site=='MP')
d2 <- mmjul %>% filter(datatype=='mm' & datasource=='insitu' & site=='PP')
instmean <- rowMeans(cbind(d1$temp, d2$temp), na.rm=TRUE)
rm(d1,d2)

# 2nd: SST data
d1 <- mmjul %>% filter(datatype=='mm' & datasource=='SST' & site=='MP')
d2 <- mmjul %>% filter(datatype=='mm' & datasource=='SST' & site=='PP')
SSTmean <- rowMeans(cbind(d1$temp, d2$temp), na.rm=TRUE)
rm(d1)

# 3rd: make offset table: SST & in situ difference by julian day
offset <- SSTmean - instmean
offsetm <- data.frame(cbind(offset,d2$dayindex))
colnames(offsetm) <- c("offset","dayindex")
str(offsetm)
rm(offset,d2)
d <- data.frame(cbind(SSTmean,instmean))
d$dayindex <- seq(from=1,to=length(d[,1]))
d <- d %>% gather("datatype","value",1:2)
head(d)
ggplot() +
  geom_line(data=d, aes(x=dayindex,y=value,color=datatype))

# 4th: adjust the raw SST by the offset, offset is per julian day
SSTo <- left_join(SST,offsetm,by="dayindex") 
SSTo <- SSTo[!is.na(SSTo$offset),]
SSTo$SSTaj <- SSTo$temp - SSTo$offset

```





```{r error=FALSE, include=FALSE, message=FALSE, warning=FALSE }
## Rolling 4 day window of stress accumulation 

#This model calculates the linear regression coefficients on 4 days of temp data. And then estimates the 'expected' temperature on the 4th day. This 'expected' temperature is one data point in the smoothed function. The difference between the 'expected' temperature and 'experienced' is a measure of stress. 

calc.stress <- function(d,wtsp){
  
  # d = data frame with at least these columns: dailyT,date
    # wtsp = vector of weights that say which days should count more in regression
  # the length(wtsp) is the window length

  # note:
  # add input parm that specifies the type of data (sst or insitu)
  # so i can add column to output df
  window <- length(wtsp)
 
  # -- Calculate Expected T in rolling window -- #
  
  expectedT <- rep(NA,length=length(d$date)-window)
  date <- rep(NA,length=length(expectedT))
  expT <- data.frame(cbind(expectedT,date))
  rm(expectedT,date)
  
  for(i in 1:(length(d$date)-4)){ #step through dates
    
    # subset temp data to window
    tsub <- d[i:(i+window-1),]
    tsub$dayseq <- c(1:window)
    tsub$wtsp <- wtsp
    
    # test by plotting (for troubleshooting)
    #plot(x=tsub$dayseq,y=tsub$dailyT,main="4 data points, 100 in blue")
    #abline(lm(dailyT~dayseq,data=tsub) )
    #abline(lm(dailyT~dayseq,data=texp),col="blue" )
    
    # add more temp observations in proportion to their weight
    texp <- tsub[rep(row.names(tsub),tsub$wtsp),1:ncol(tsub)-1]
    
    # calc regression, predict 'expected T' on last day in window
    expT$expectedT[i] <- predict(lm(data=texp,dailyT~dayseq),
                                 list(dayseq=tsub$dayseq[window]))
    expT$date[i] <- as.character(tsub$date[window])
  }
  expT$date <- as.Date(expT$date,format="%Y-%m-%d")
  d2 <- left_join(d,expT,by="date")
  d2$Dstress <- d2$dailyT - d2$expectedT
  rm(i,tsub,texp)
  return(d2)
  
}

```



#### 1) Plot the raw temperature at each location.

*In situ* daily temperature is randomly chosen from 10 minute temperature recordings.

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=6}

# plot 1: raw data at each site
# these two lines calculate the mean daily T
#d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
#d2 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
# these two lines randomly selects one temperature recording from each day
d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% sample_n(1)
d2 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% sample_n(1)
names(d1)[names(d1) == "Temperature"] <- "dailyT"
names(d2)[names(d2) == "Temperature"] <- "dailyT"  
 
# both sites together
p1 <- ggplot() +
  geom_line(data=d1, aes(x=date,y=dailyT,color="MP_insitu")) +
  geom_smooth(data=d1, aes(x=date,y=dailyT,color="MP_insitu"),se=FALSE) +
  
  geom_line(data=d2,aes(x=date,y=dailyT,color="PP_insitu")) +
  geom_smooth(data=d2,aes(x=date,y=dailyT,color="PP_insitu"),se=FALSE) +
  
  geom_line(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst")) + 
  geom_smooth(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst"),se=FALSE) + 
  
  geom_line(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst")) +
  geom_smooth(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst"),se=FALSE) +
  
  ggtitle("Raw in situ and SST data at Morro Prieto & Punta Prieta") +
  ylab("daily T (randomly selected observation per day)")

# Morro Prieto
p2 <- ggplot() +
  geom_line(data=d1, aes(x=date,y=dailyT,color="MP_insitu")) +
  geom_smooth(data=d1, aes(x=date,y=dailyT,color="MP_insitu"),se=FALSE) +
  
  geom_line(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst")) +
  geom_smooth(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst"),se=FALSE) +
  ylim(c(11,27)) + ggtitle("a) Morro Prieto ") +
  ylab("daily T\n(random observation per day)")

# Punta Prieta
p3 <- ggplot() +
 
  geom_line(data=d2,aes(x=date,y=dailyT,color="PP_insitu")) +
  geom_smooth(data=d2,aes(x=date,y=dailyT,color="PP_insitu"),se=FALSE) +
  
  geom_line(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst")) + 
  geom_smooth(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst"),se=FALSE) + 
  ylim(c(11,27)) + ggtitle("b) Punta Prieta") +
  ylab("daily T\n(random observation per day)")

grid.newpage()
grid.draw(rbind(ggplotGrob(p2), ggplotGrob(p3), size = "last"))
rm(p1,p2,p3,d1,d2)

```

*In situ* daily temperature is average of 10 minute temperature recordings.

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=6}

# plot 1: raw data at each site
# these two lines calculate the mean daily T
d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d2 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))

# both sites together
p1 <- ggplot() +
  geom_line(data=d1, aes(x=date,y=dailyT,color="MP_insitu")) +
  geom_smooth(data=d1, aes(x=date,y=dailyT,color="MP_insitu"),se=FALSE) +
  
  geom_line(data=d2,aes(x=date,y=dailyT,color="PP_insitu")) +
  geom_smooth(data=d2,aes(x=date,y=dailyT,color="PP_insitu"),se=FALSE) +
  
  geom_line(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst")) + 
  geom_smooth(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst"),se=FALSE) + 
  
  geom_line(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst")) +
  geom_smooth(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst"),se=FALSE) +
  
  ggtitle("Raw in situ and SST data at Morro Prieto & Punta Prieta") +
  ylab("daily T (randomly selected observation per day)")

# Morro Prieto
p2 <- ggplot() +
  geom_line(data=d1, aes(x=date,y=dailyT,color="MP_insitu")) +
  geom_smooth(data=d1, aes(x=date,y=dailyT,color="MP_insitu"),se=FALSE) +
  
  geom_line(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst")) +
  geom_smooth(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst"),se=FALSE) +
  ylim(c(11,27)) + ggtitle("a) Morro Prieto ") +
  ylab("daily T\n(mean temperature)")

# Punta Prieta
p3 <- ggplot() +
 
  geom_line(data=d2,aes(x=date,y=dailyT,color="PP_insitu")) +
  geom_smooth(data=d2,aes(x=date,y=dailyT,color="PP_insitu"),se=FALSE) +
  
  geom_line(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst")) + 
  geom_smooth(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst"),se=FALSE) + 
  ylim(c(11,27)) + ggtitle("b) Punta Prieta") +
  ylab("daily T\n(mean temperature)")

grid.newpage()
grid.draw(rbind(ggplotGrob(p2), ggplotGrob(p3), size = "last"))
rm(p1,p2,p3,d1,d2)

```


#### 2) Skip 

Old stuff: Plot the average temperature per Julian day. The daily offset is calculated from this plot: SSTmean - instmean (for each julian day). The daily offset is substracted from raw daily SST (see fig 3)

```{r eval=FALSE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}

p1 <- ggplot()+
  #geom_line(data=in_situ,aes(x=dayindex,y=Temperature,color=Site)) +
  geom_point(data=mmjul[mmjul$datatype=="dailyT",],
            aes(x=dayindex,y=temp,color=sitesr)) + 
  geom_line(data=mmjul[mmjul$datatype=="mm",],
            aes(x=dayindex,y=temp,color=sitesr)) + 
  ggtitle("Julian day averages\nUse this plot to calculate daily offset on SST\nOffset = (moving mean SST per julian day) - (moving mean insitu per julian day)") 

# plot overall average of in situ and SST temp
d <- data.frame(cbind(SSTmean,instmean))
d$dayindex <- seq(from=1,to=length(d[,1]))
d <- d %>% gather("datatype","value",1:2)
p2 <- ggplot() +
  geom_line(data=d, aes(x=dayindex,y=value,color=datatype)) +
  xlab("Julian day") + ylab("Temp")

p2
rm(p1,p2,d)
```

#### 3) Skip

Old stuff: Compare raw SST with the adjusted SST (SST-offset) 

```{r eval=FALSE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
#plot only dates between 

ggplot() +
   # raw SST
   geom_line(data=SSTo[SSTo$date > as.Date("2015-09-12") & SSTo$date < as.Date("2018-08-17"),],
              aes(x=date,y=temp),color="grey") +
  # ajusted SST
  geom_line(data=SSTo[SSTo$date > as.Date("2015-09-12") & SSTo$date < as.Date("2018-08-17"),],
             aes(x=date,y=SSTaj,color=datatype)) +
  
  ggtitle("Raw SST (grey) & adjusted SST")
```

#### 4) Skip

Overlay the adjusted SST with *in situ* time series -- do they match?

The goal is to have the adjusted SST time series (here, the blue and green lines) overlap the *in situ* time series as much as possible (the red and purple lines). 

Does the daily SST offset work?

* Sort of. The peaks and troughs appear to line up more closely than the raw SST data

* Mean temperature per day in the adjusted SST time series doesn't overlap the mean *in situ* throughout the year (this is bad)

* In some years there is better overlap compared to other years. 


##### 4 (a,b) Skip

Overlap between adjusted SST and in situ lines is close

Offset in SST is based on an average of in situ data at both Morro Prieto and Punta Prieta. 

```{r eval=FALSE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=6}

# both sites
p1 <- ggplot() +
  
  # raw SST
  # geom_line(data=SSTo[SSTo$date > as.Date("2015-09-12") & SSTo$date < as.Date("2018-08-17"),],
  #            aes(x=date,y=temp),color="grey") +
  # 
  #raw in situ
  geom_point(data=in_situ,
              aes(x=date,y=Temperature,color=Site)) +
  # ajusted SST
  geom_line(data=SSTo[SSTo$date > as.Date("2015-09-12") & SSTo$date < as.Date("2018-08-17"),],
             aes(x=date,y=SSTaj,color=datatype)) +
  # raw in situ
  geom_smooth(data=in_situ,
              aes(x=date,y=Temperature,color=Site)) +
  ggtitle("SST is adjusted with offset\n Adjusted SST and in situ lines overlap, but not that great")

# Morro Prieto 

p2 <- ggplot() +
  
  #raw in situ
  geom_line(data=in_situ[in_situ$Site=="MorroPrieto",],
              aes(x=date,y=Temperature,color=Site)) +
  #ajusted SST
  geom_line(data=SSTo[SSTo$date > as.Date("2015-09-12") & 
                        SSTo$date < as.Date("2018-08-17") & 
                        SSTo$datatype=="MP_SST",],
             aes(x=date,y=SSTaj,color=datatype)) +
  ylim(c(12,26)) + ggtitle("a) Morro Prieto")

p3 <- ggplot() +
  
  #raw in situ
  geom_line(data=in_situ[in_situ$Site=="PuntaPrieta",],
              aes(x=date,y=Temperature,color=Site)) +
  #ajusted SST
  geom_line(data=SSTo[SSTo$date > as.Date("2015-09-12") & 
                        SSTo$date < as.Date("2018-08-17") & 
                        SSTo$datatype=="PP_SST",],
             aes(x=date,y=SSTaj,color=datatype)) + 
  ylim(c(12,26)) + ggtitle("b) Punta Prieta")

grid.newpage()
grid.draw(rbind(ggplotGrob(p2), ggplotGrob(p3), size = "last"))
rm(p1,p2,p3)

```


#### 5) Skip

Old stuff: Apply stress model to adjusted SST & *in situ* time series

Daily stress in 2016 using *in situ* time series and adjusted SST time series 

```{r eval=FALSE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# assign weights: abalone acclimiatize to the temperature today most, and 
# the temperature on the earliest days least
#wts <- c(0.1,0.2,0.4,0.8)
wts <- c(0.1,0.4,0.8)
wts <- wts/sum(wts)
wtsp <- round(wts*100,digits=0)

# Calculate daily stress for:

# MP in situ 
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
d1 <- dd1
MP_inst_stress <- calc.stress(d=d1,wtsp=wtsp)
MP_inst_stress$datatypesr <- rep("MP_insitu",length=length(MP_inst_stress[,1]))
MP_inst_stress <- MP_inst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1,dd1)

# -- SKIP -- MP SST adjusted
d1 <- SSTo[SSTo$date > as.Date("2015-09-12") & SSTo$date < as.Date("2018-08-17") & SSTo$datatype=="MP_SST",]
names(d1)[names(d1)=="SSTaj"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
MP_sst_stress <- calc.stress(d=d1,wtsp=wtsp)
MP_sst_stress$datatypesr <- rep("MP_SST",length=length(MP_sst_stress[,1]))
MP_sst_stress <- MP_sst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1)

# PP in situ
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
d1 <- dd1
PP_inst_stress <- calc.stress(d=d1,wtsp=wtsp)
PP_inst_stress$datatypesr <- rep("PP_insitu",length=length(PP_inst_stress[,1]))
PP_inst_stress <- PP_inst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1,dd1)

# -- SKIP -- PP SST adjusted
d1 <- SSTo[SSTo$date > as.Date("2015-09-12") & SSTo$date < as.Date("2018-08-17") & SSTo$datatype=="PP_SST",]
names(d1)[names(d1)=="SSTaj"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
PP_sst_stress <- calc.stress(d=d1,wtsp=wtsp)
PP_sst_stress$datatypesr <- rep("PP_SST",length=length(PP_sst_stress[,1]))
PP_sst_stress <- PP_sst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1)

# subset SST to only dates overlapping with in situ
d1 <- MP_sst_stress[MP_sst_stress$date %in% MP_inst_stress$date,]
d1 <- bind_rows(d1,MP_inst_stress)

d2 <- PP_sst_stress[PP_sst_stress$date %in% PP_inst_stress$date,]
d2 <- bind_rows(d2,PP_inst_stress)
```


```{r eval=FALSE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=6}
# plot stress for SST & in situ
p1 <- ggplot(data=d1[d1$date > as.Date("2016-01-01") & d1$date < as.Date("2016-06-29"),]) +
  geom_line(aes(x=date,y=Dstress,color=datatypesr)) +
  xlab("date (2016 only)") + ggtitle("a) Morro Prieto")

p2 <- ggplot(data=d2[d2$date > as.Date("2016-01-01") & d1$date < as.Date("2016-06-29"),]) +
  geom_line(aes(x=date,y=Dstress,color=datatypesr)) +
  xlab("date (2016 only)") + ggtitle("b) Punta Prieta")

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), size = "last"))
rm(p1,p2,d1,d2)

```



## Apply the stress model on raw SST and *in situ* time series


```{r eval=FALSE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
#### As before, here are the raw temperatures at each location
# plot 1: raw data at each site
d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d2 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))

ggplot() +
  geom_line(data=d1, aes(x=date,y=dailyT,color="MP_insitu")) +
  geom_smooth(data=d1, aes(x=date,y=dailyT,color="MP_insitu"),se=FALSE) +
  
  geom_line(data=d2,aes(x=date,y=dailyT,color="PP_insitu")) +
  geom_smooth(data=d2,aes(x=date,y=dailyT,color="PP_insitu"),se=FALSE) +
  
  geom_line(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst")) + 
  geom_smooth(data=SST[SST$datatype=="PP_SST" & 
SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="PP_sst"),se=FALSE) + 
  
  geom_line(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst")) +
  geom_smooth(data=SST[SST$datatype=="MP_SST" & SST$date > as.Date("2015-09-12") & SST$date < as.Date("2018-08-17"),],aes(x=date,y=temp,color="MP_sst"),se=FALSE) +
  
  ggtitle("Raw in situ and SST data at Morro Prieto & Punta Prieta")
```

We run the stress model on raw time series. *In situ* daily time series is not the average daily temperature, daily bottom temperature is randomly selected from the 10 minute data. We did this to have a more direct comparison with daily SST observations. 


```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# assign weights: abalone acclimiatize to the temperature today most, and 
# the temperature on the earliest days least
#wts <- c(0.1,0.2,0.4,0.8)
wts <- c(0.1,0.4,0.8)
wts <- wts/sum(wts)
wtsp <- round(wts*100,digits=0)

# Calculate daily stress for:

# MP in situ 
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
d1 <- dd1
MP_inst_stress <- calc.stress(d=d1,wtsp=wtsp)
MP_inst_stress$datatypesr <- rep("MP_insitu",length=length(MP_inst_stress[,1]))
MP_inst_stress <- MP_inst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1,dd1)

# MP SST
d1 <- SST[SST$datatype == "MP_SST",]
names(d1)[names(d1)=="temp"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
MP_sst_stress <- calc.stress(d=d1,wtsp=wtsp)
MP_sst_stress$datatypesr <- rep("MP_SST",length=length(MP_sst_stress[,1]))
MP_sst_stress <- MP_sst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1)

# PP in situ
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
d1 <- dd1
PP_inst_stress <- calc.stress(d=d1,wtsp=wtsp)
PP_inst_stress$datatypesr <- rep("PP_insitu",length=length(PP_inst_stress[,1]))
PP_inst_stress <- PP_inst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1,dd1)


# PP SST
d1 <- SST[SST$datatype == "PP_SST",]
names(d1)[names(d1)=="temp"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
PP_sst_stress <- calc.stress(d=d1,wtsp=wtsp)
PP_sst_stress$datatypesr <- rep("PP_SST",length=length(PP_sst_stress[,1]))
PP_sst_stress <- PP_sst_stress %>% select(date,dailyT,expectedT,Dstress,datatypesr)
rm(d1)

# subset SST to only dates overlapping with in situ
d1 <- MP_sst_stress[MP_sst_stress$date %in% MP_inst_stress$date,]
d1 <- bind_rows(d1,MP_inst_stress)

d2 <- PP_sst_stress[PP_sst_stress$date %in% PP_inst_stress$date,]
d2 <- bind_rows(d2,PP_inst_stress)
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=6}
# plot stress for SST & in situ
p1 <- ggplot(data=d1[d1$date > as.Date("2016-04-01") & d1$date < as.Date("2016-09-29"),]) +
  geom_line(aes(x=date,y=Dstress,color=datatypesr)) +
  xlab("date (2016 only)") + ggtitle("a) Morro Prieto")+
  ylab("Dstress (expected T - experienced T)")

p2 <- ggplot(data=d2[d2$date > as.Date("2016-04-01") & d2$date < as.Date("2016-09-29"),]) +
  geom_line(aes(x=date,y=Dstress,color=datatypesr)) +
  xlab("date (2016 only)") + ggtitle("b) Punta Prieta")+
  ylab("Dstress (expected T - experienced T)")

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), size = "last"))
rm(p1,p2)


```

#### Plot stress time series for entire year (2016)

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=8}
# plot stress for SST & in situ
p1 <- ggplot(data=d1[d1$date > as.Date("2016-01-01") & d1$date < as.Date("2016-12-29"),]) +
  geom_line(aes(x=date,y=Dstress,color=datatypesr)) +
  xlab("date (2016 only)") + ggtitle("a) Morro Prieto") +
  ylab("Dstress (expected T - experienced T)")

p2 <- ggplot(data=d2[d2$date > as.Date("2016-01-01") & d2$date < as.Date("2016-12-29"),]) +
  geom_line(aes(x=date,y=Dstress,color=datatypesr)) +
  xlab("date (2016 only)") + ggtitle("b) Punta Prieta")+
  ylab("Dstress (expected T - experienced T)")

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), size = "last"))
rm(p1,p2,d1,d2)


```


## Time lags between SST and in situ

For one site, plot SST at time *t* on x-axis and in situ temperature at *t+n* day. Here *n* is a number of days into the future from day *t*. 

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# prep for making lag df: randomly select dailyT from in situ

# MP in situ 
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
mp_insitu <- d1
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
mp_insitu <- dd1
rm(d1,dd1)
# MP SST
d1 <- SST[SST$datatype == "MP_SST",]
names(d1)[names(d1)=="temp"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
mp_sst <- d1
mp_sst <- mp_sst[mp_sst$date %in% mp_insitu$date,]
rm(d1)

# PP in situ 
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
mp_insitu <- d1
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
pp_insitu <- dd1
rm(d1,dd1)
# PP SST
d1 <- SST[SST$datatype == "PP_SST",]
names(d1)[names(d1)=="temp"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
pp_sst <- d1
pp_sst <- pp_sst[pp_sst$date %in% pp_insitu$date,]
rm(d1)


# days into the future from day t
n = 30

# fill in this data frame -- MP
columns <-  c("datet","datetn","Tt_sst","Ttn_insitu","site")
d <- data.frame(matrix(NA,nrow=(length(unique(in_situ$date))-n),ncol=length(columns)))
names(d) <- columns
rm(columns)
# fill in dates & T -- MP
d$datet <- mp_insitu$date[1:(length(mp_insitu$date)-n)]
d$datetn <- mp_insitu$date[(n+1):length(mp_insitu$date)]
d$Tt_sst <- mp_sst$dailyT[1:(length(mp_sst$dailyT)-n)] #SST at t
d$Ttn_insitu <- mp_insitu$dailyT[(n+1):length(mp_insitu$dailyT)] #bottom at t+n
d$site <- rep("MP")
mpd <- d
rm(d)

# fill in this data frame -- PP
columns <-  c("datet","datetn","Tt_sst","Ttn_insitu","site")
d <- data.frame(matrix(NA,nrow=(length(unique(in_situ$date))-n),ncol=length(columns)))
names(d) <- columns

rm(columns)
# fill in dates & T -- PP
d$datet <- pp_insitu$date[1:(length(pp_insitu$date)-n)]
d$datetn <- pp_insitu$date[(n+1):length(pp_insitu$date)]
d$Tt_sst <- pp_sst$dailyT[1:(length(pp_insitu$dailyT)-n)] #SST at t
d$Ttn_insitu <- pp_insitu$dailyT[(n+1):length(pp_sst$dailyT)] #bottom at t+n
d$site <- rep("PP")
ppd <- d
rm(d)

```

### 1) `r n` day time lag

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=6}
# plot SST at t vs insitu at t+n
p1 <- ggplot(data=mpd) +
  geom_point(aes(x=Tt_sst,y=Ttn_insitu)) +
  xlab("SST (t)") + 
  ylab(paste("bottom temperature (t+",n,")",sep="")) +
  ggtitle(paste("a) Morro Prieto"))

p2 <- ggplot(data=ppd) + 
  geom_point(aes(x=Tt_sst,y=Ttn_insitu)) +
  xlab("SST (t)") + 
  ylab(paste("bottom temperature (t+",n,")",sep="")) +
  ggtitle("b) Punta Prieta")

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), size = "last"))
rm(p1,p2,d,dd)

```



```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# prep for making lag df: randomly select dailyT from in situ

# MP in situ 
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
mp_insitu <- d1
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "MorroPrieto",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
mp_insitu <- dd1
rm(d1,dd1)
# MP SST
d1 <- SST[SST$datatype == "MP_SST",]
names(d1)[names(d1)=="temp"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
mp_sst <- d1
mp_sst <- mp_sst[mp_sst$date %in% mp_insitu$date,]
rm(d1)

# PP in situ 
# -- these two lines finds the average daily T -- #
d1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% summarize(dailyT = mean(Temperature,na.rm=TRUE))
d1 <- left_join(d1,datedf,by='date')
mp_insitu <- d1
# -- these two lines randomly selects one temperature recording from each day -- #
dd1 <- in_situ[in_situ$Site == "PuntaPrieta",] %>% group_by(date) %>% sample_n(1)
names(dd1)[names(dd1) == "Temperature"] <- "dailyT"
dd1 <- dd1 %>% select(names(d1))
pp_insitu <- dd1
rm(d1,dd1)
# PP SST
d1 <- SST[SST$datatype == "PP_SST",]
names(d1)[names(d1)=="temp"] <- "dailyT"
d1 <- left_join(d1,datedf,by='date')
pp_sst <- d1
pp_sst <- pp_sst[pp_sst$date %in% pp_insitu$date,]
rm(d1)


# days into the future from day t
n = 45

# fill in this data frame -- MP
columns <-  c("datet","datetn","Tt_sst","Ttn_insitu","site")
d <- data.frame(matrix(NA,nrow=(length(unique(in_situ$date))-n),ncol=length(columns)))
names(d) <- columns
rm(columns)
# fill in dates & T -- MP
d$datet <- mp_insitu$date[1:(length(mp_insitu$date)-n)]
d$datetn <- mp_insitu$date[(n+1):length(mp_insitu$date)]
d$Tt_sst <- mp_sst$dailyT[1:(length(mp_sst$dailyT)-n)] #SST at t
d$Ttn_insitu <- mp_insitu$dailyT[(n+1):length(mp_insitu$dailyT)] #bottom at t+n
d$site <- rep("MP")
mpd <- d
rm(d)

# fill in this data frame -- PP
columns <-  c("datet","datetn","Tt_sst","Ttn_insitu","site")
d <- data.frame(matrix(NA,nrow=(length(unique(in_situ$date))-n),ncol=length(columns)))
names(d) <- columns

rm(columns)
# fill in dates & T -- PP
d$datet <- pp_insitu$date[1:(length(pp_insitu$date)-n)]
d$datetn <- pp_insitu$date[(n+1):length(pp_insitu$date)]
d$Tt_sst <- pp_sst$dailyT[1:(length(pp_insitu$dailyT)-n)] #SST at t
d$Ttn_insitu <- pp_insitu$dailyT[(n+1):length(pp_sst$dailyT)] #bottom at t+n
d$site <- rep("PP")
ppd <- d
rm(d)

```

### 2) `r n` day time lag

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=6}
# plot SST at t vs insitu at t+n
p1 <- ggplot(data=mpd) +
  geom_point(aes(x=Tt_sst,y=Ttn_insitu)) +
  xlab("SST (t)") + 
  ylab(paste("bottom temperature (t+",n,")",sep="")) +
  ggtitle(paste("a) Morro Prieto"))

p2 <- ggplot(data=ppd) + 
  geom_point(aes(x=Tt_sst,y=Ttn_insitu)) +
  xlab("SST (t)") + 
  ylab(paste("bottom temperature (t+",n,")",sep="")) +
  ggtitle("b) Punta Prieta")

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), size = "last"))
rm(p1,p2,d,dd)

```
