---
title: "Optimize time lag & moving average window:\n AR model & temperature spectra"
author: "Mikaela M. Provost"
date: "8/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plan
0. Read in data
1. For each site, plot spectra of raw data. 10 min bottom, daily grab bottom, SST daily
2. For each site, plot raw data. 10 min bottom, daily grab bottom, SST daily
3. For each site, plot mSST (SST + offset + deviations in daily grab bottom) & daily grab bottom
4. For each site, plot spectrum of mSST & daily grab bottom
5. For each site, plot spectrum of mSST with different weights & lags


6. Use spectrum to generate noise(?) See Will's code. 


```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(grid)
library(gridExtra)

```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# -- Read in data -- #

# read in data - bottom
d <- read.csv(file="C:/Users/Mikaela/Documents/GitHub/natividad/data/001_FEDECOOP_MS_DATA.csv",stringsAsFactors = FALSE)
d$Date <- as.Date(d$UTC_DateTime,format="%Y-%m-%d") # create date-only column
bottomfede <- d %>% group_by(Site,Date) %>% sample_n(1) %>% select(Site,Temperature,Date)# randomly select temp per day
d1 <- read.csv(file="C:/Users/Mikaela/Documents/GitHub/natividad/data/001_IslaNatividad_MorroPrietoPuntaPrieta_alldata_trimmed.csv",stringsAsFactors = FALSE)
d1$Date <- as.Date(d1$UTC_Date_Time,format="%Y-%m-%d")
bottommppp <- d1 %>% 
  filter(Site %in% c('MorroPrieto','PuntaPrieta'), Instrument_Type == 'MiniDOT') %>% 
  group_by(Site,Date) %>% 
  sample_n(1) %>% 
  select(Site,Temperature,Date)
bottomall <- rbind(bottomfede,bottommppp)
names(bottomall)[names(bottomall)=="Temperature"] <- "temp_bottom"
rm(d1,d,bottomfede,bottommppp)

# read in data - SST
m <- read.csv(file= "C:/Users/Mikaela/Documents/GitHub/natividad/data/morroprieto_temp.csv",stringsAsFactors = FALSE)
p <- read.csv(file= "C:/Users/Mikaela/Documents/GitHub/natividad/data/puntaprieta_temp.csv",stringsAsFactors = FALSE)
m$Site <- rep("Morro_Prieto")
p$Site <- rep("Punta_Prieta")
p$datetime <- as.Date(p$datetime,format="%m/%d/%Y") #more fun with dates
m$datetime <- as.Date(m$datetime,format="%m/%d/%Y") #I hate dates
file_names <- list.files(path = "C:/Users/Mikaela/Documents/GitHub/natividad/data/sstfedecoop",pattern = "csv") 
file_names <- unlist(strsplit(file_names,split="_temp.csv"), use.names=FALSE)
filesL <- as.list(rep(NA,length=length(file_names)))
names(filesL) <- file_names
for(f in 1:length(file_names)){
  filesL[[f]] <- read.csv(paste("C:/Users/Mikaela/Documents/GitHub/natividad/data/sstfedecoop/",
                file_names[f],"_temp.csv",sep=""),stringsAsFactors = FALSE)} #end loop
SSTfede <- bind_rows(filesL,.id="Site")
SSTfede$datetime <- as.Date(SSTfede$datetime,format="%d-%b-%Y") #date problems
SSTall <- bind_rows(p,m,SSTfede)
SSTall <- SSTall %>% select(datetime,temp,Site,flag)
names(SSTall)[names(SSTall)=="datetime"] <- "Date"
names(SSTall)[names(SSTall)=="temp"] <- "temp_sst"
rm(f,m,p,SSTfede,filesL,file_names)

# standardize site location names
# export the site names from each dataset, manually rearrange in excel and read back in
# bsites <- unique(bottomall$Site)
# Ssites <- unique(SSTall$Site) 
# siteskey <- cbind(bsites,Ssites)
# write.csv(siteskey,file="C:/Users/Mikaela/Documents/GitHub/natividad/data/siteskey.csv")
# rm(bsites,Ssites,siteskey)
sitekey <- read.csv(file="C:/Users/Mikaela/Documents/GitHub/natividad/data/siteskey_fixed.csv",stringsAsFactors = F)
sitekey <- sitekey[complete.cases(sitekey),]
SSTall$bsite <- sitekey[match(SSTall$Site,sitekey$Ssites),"bsites"]
bottomall$Ssite <- sitekey[match(bottomall$Site,sitekey$bsites),"Ssites"] 

# keep same dates
combo <- left_join(SSTall,bottomall,by=c("Date"="Date","Site"="Ssite"))
combo <- combo[complete.cases(combo),]
combo <- combo %>% select(Date,temp_sst,Site,temp_bottom)
rm(sitekey,bottomall,SSTall)


# (0) for each site
temp_pL <- as.list(rep(NA,length=length(unique(combo$Site))))
r2_pL <- as.list(rep(NA,length=length(unique(combo$Site))))
best_lags <- as.data.frame(matrix(rep(NA,length=length(unique(combo$Site))*2),
                                        nrow=length(unique(combo$Site)),
                                        ncol=2))
names(best_lags) <- c("dayslag","site")
lags = seq(from=15,to=45, by=1)

for(s in 1:length(unique(combo$Site))){ # for each site

# (1) apply large moving average window to capture seasal variation
  dd <- combo[combo$Site==unique(combo$Site)[s],]

  # plot raw -- bottom
  b <- ggplot(data=dd,aes(x=Date,y=temp_bottom)) +
    geom_line() + xlab("date") + ylab("T") + 
    geom_smooth(se=FALSE,n=length(dd$Date)) 
  # get data for large moving average window -- bottom
  b1 <- ggplot_build(b)$data[[2]]
  
  # plot raw -- bottom
  t <- ggplot(data=dd,aes(x=Date,y=temp_sst)) +
    geom_line() + xlab("date") + ylab("T") +  
    geom_smooth(se=FALSE,n=length(dd$Date)) 
  # get data for large moving average window -- bottom
  t1 <- ggplot_build(t)$data[[2]]
  
  # prep to find best lag time
  r2 <- rep(NA,length=length(lags))
  # search for time lag
  for(L in 1:length(lags)){
    # make empty d for lag temps
    columns <-  c("datet","datetn","Tt_sst","Ttn_bottom","site")
    d <- data.frame(matrix(NA,nrow=(length(t1$y)-lags[L]),ncol=length(columns)))
    names(d) <- columns
    rm(columns)
    # fill in d
    d$datet <- dd$Date[1:(length(dd$Date)-lags[L])]
    d$datetn <- dd$Date[(lags[L]+1):length(dd$Date)]
    d$Tt_sst <- t1$y[1:(length(dd$temp_sst)-lags[L])] #SST at t
    d$Ttn_bottom <- b1$y[(lags[L]+1):length(b1$y)] #bottom at t+n
    d$site <- rep(dd$Ste[1])
    #plot(x=d$Tt_sst,y=d$Ttn_bottom)
    r2[L] <- summary(lm(d$Ttn_bottom ~ d$Tt_sst))$r.squared
  }
  lagdf <- as.data.frame(cbind(lags,r2))
  #plot(x=lagdf$lags,y=lagdf$r2)
  best_lags$dayslag[s] <- lagdf[lagdf$r2==max(lagdf$r2),]$lags
  best_lags$site[s] <- dd$Site[1]
  
  # save temp plots with window info
  temp_pL[[s]] <- ggplot(data=dd) + 
    geom_line(aes(x=Date,y=temp_sst,color="SST")) +
    geom_smooth(aes(x=Date,y=temp_sst,color="SST"),se=F) +
    geom_line(aes(x=Date,y=temp_bottom,color="bottom")) +
    geom_smooth(aes(x=Date,y=temp_bottom,color="bottom"),se=F) +
    ggtitle(paste(dd$Site[1])) + 
    theme(axis.text.x = element_text(angle = 45,hjust=1),legend.position="none") +
    ylab("") + xlab("")
  
  # time lag vs r^2 
  r2_pL[[s]] <- ggplot(data=lagdf,aes(x=lags,y=r2)) + 
    geom_point() +
    xlab("") + ylab("") + 
    ggtitle(paste(dd$Site[1]))
  
}
rm(b,b1,t,t1,d,dd,lagdf,L,s,r2,lags)
# tiff(file='C:/Users/Mikaela/Documents/GitHub/natividad/plots/sst_bottom/raw_temps.tiff', 
#      units="in", width=7, height=14, res=300) 
# do.call(grid.arrange,c(temp_pL,ncol=2,left="",bottom=""))
# dev.off()
# 
# tiff(file='C:/Users/Mikaela/Documents/GitHub/natividad/plots/sst_bottom/lags_r2.tiff', 
#      units="in", width=7, height=14, res=300) 
# do.call(grid.arrange,c(r2_pL,ncol=2,left="",bottom=""))
# dev.off()


```

### Fig 1. Raw SST and bottom temperature at each site with large moving window

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=14, fig.width=8}
do.call(grid.arrange,c(temp_pL,ncol=2,left="Tempertaure",bottom=""))
```

### Fig 2.  For each site, plot spectra of raw data. 10 min bottom, daily grab bottom, SST daily

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
ma <- function(x, n = 2){stats::filter(x, rep(1 / n, n), sides = 2)} #function to calc moving average



# spectrum for 10 min data 
d1 <- read.csv(file="C:/Users/Mikaela/Documents/GitHub/natividad/data/001_IslaNatividad_MorroPrietoPuntaPrieta_alldata_trimmed.csv",stringsAsFactors = FALSE)
d1$Date <- as.Date(d1$UTC_Date_Time,format="%Y-%m-%d")
bottommppp <- d1 %>% 
  filter(Site %in% c('MorroPrieto','PuntaPrieta'), Instrument_Type == 'MiniDOT') %>% 
  select(Site,Temperature,Date,UTC_Date_Time,Year)
bottom10 <- bottommppp
rm(d1,bottommppp)

# plot spectrum of raw 10 bottom
span.multiplier = 1

tmp <- ceiling(sqrt(length(1:length(bottom10[bottom10$Site=="MorroPrieto",]$Temperature)))) 
if (tmp %% 2 == 0) {m <- tmp+1} else {m <- tmp} #make it odd, if the square root is even
m = m * span.multiplier

# sampling interval of 10 min data
sampleinterval <- 10/1440 #10 min/total min in day = observation interval in terms of days
sp1 <- spec.pgram(bottom10[bottom10$Site=="MorroPrieto",]$Temperature,
                  spans=NULL,taper=0.1,plot = TRUE,demean = TRUE)
spx <- sp1$freq/sampleinterval
spy <- sp1$spec*sp1$freq

plot(x=sp1$freq,y=log10(sp1$spec),xlab="frequency",ylab="spectral density",type="l")

p1 <- ggplot() + geom_line(aes(x=sp1$freq,y=log10(sp1$spec))) +
    ggtitle("raw SST ") + xlab("freq (units = 1/10 min observation)")
  
p2 <- ggplot() + geom_line(aes(x=(spx),y=log10(spy))) +
    ggtitle("raw SST ") + xlab("cycles per day")

rawspec <- as.list(rep(NA,length=length(unique(combo$Site))))
names(rawspec) <- unique(combo$Site)

for(s in 1:length(rawspec)){
  # subset to a site
  dd <- combo[combo$Site==unique(combo$Site)[s],]
    # square root time series length
  tmp <- ceiling(sqrt(length(1:length(dd$temp_bottom)))) #sq root of timeseries lgth, rounded
  if (tmp %% 2 == 0) {m <- tmp+1} else {m <- tmp} #make it odd, if the square root is even
  m = m * span.multiplier
  
  # sst raw spec
  sp1 <- spec.pgram(dd$temp_sst,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p1 <- ggplot() + geom_line(aes(x=sp1$freq,y=log10(sp1$spec))) +
    ggtitle("raw SST ") + xlab("freq")
  
  # bottom raw spec
  sp2 <- spec.pgram(dd$temp_bottom,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p2 <- ggplot() + geom_line(aes(x=sp2$freq,y=log10(sp2$spec))) +
    ggtitle("raw bottom ") + xlab("freq")
  
  
  # -- later -- #
  # sst smooth spec
  sstS <- ma(dd$temp_sst,n=6)
  sstS <- sstS[!is.na(sstS)]
  sp3 <- spec.pgram(sstS,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p3 <- ggplot() + geom_line(aes(x=sp3$freq,y=log10(sp3$spec))) +
    ggtitle("smooth SST") + xlab("freq")
  #ggtitle(paste(dd$Site[1]," - smooth SST ",sep="")) + xlab("freq")
  
  # bottom smooth spec
  bottomS <- ma(dd$temp_bottom,n=6)
  bottomS <- bottomS[!is.na(bottomS)]
  sp4 <- spec.pgram(bottomS,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p4 <- ggplot() + geom_line(aes(x=sp4$freq,y=log10(sp4$spec))) +
    ggtitle("smooth bottom ") + xlab("freq")
  
  siteplots <- list(p1,p2,p3,p4)
  
  specplots[[s]] <- siteplots
  
}

```




```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=14, fig.width=8}
do.call(grid.arrange,c(r2_pL,ncol=2,left="r2 (bottom ~ SST)", bottom="Time lag (days)"))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# for each site, run through different moving average windows
windows <- c(3,4,5,6,7,8,9)
ma <- function(x, n = 2){stats::filter(x, rep(1 / n, n), sides = 2)}
pL <- as.list(rep(NA,length=length(unique(combo$Site))))


for(s in 1:length(unique(combo$Site))){
  dd <- combo[combo$Site==unique(combo$Site)[s],]
  df <- as.data.frame(matrix(rep(NA,length=length(windows)*3),
                                        nrow=length(windows),
                                        ncol=3))
  names(df) <- c("site","movingw","r2")
  for(w in 1:length(windows)){
    t <- ma(x=dd$temp_sst,n=windows[w])
    b <- ma(x=dd$temp_bottom,n=windows[w])
    d <- as.data.frame(cbind(dd$Date,t,b))
    sst <- d$t[1:(length(d$b)-best_lags[best_lags$site==dd$Site[1],]$dayslag)] #SST at t
    bottom <- d$b[(best_lags[best_lags$site==dd$Site[1],]$dayslag+1):length(d$b)] #bottom at t+n
    # r2, window, and site 
    df$r2[w] <- summary(lm(bottom ~ sst))$r.squared
    df$movingw[w] <- windows[w]
    df$site[w] <- dd$Site[1]
  }
  diffs <- diff(df$r2)
  
  pL[[s]]<- ggplot(data=df,aes(x=movingw,y=r2)) + 
    geom_point() + geom_line() + xlab("") + ylab("") + 
    ggtitle(paste(dd$Site[1]," (",windows[match(min(diffs),diffs)],")",sep="")) + 
    geom_vline(xintercept =  windows[match(min(diffs),diffs)],linetype="dashed")
}
rm(t,b,d,sst,bottom,df,s,dd,w)


```

### Fig 3. r2 between SST and bottom temperature for different moving average windows

I use site-specific time lags when calculating r2. I tested these windows: `r windows`

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=14, fig.width=8}
do.call(grid.arrange,c(pL,ncol=2,left="r2 (bottom ~ SST)",bottom="Moving average window (days)"))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
rm(windows)
ma <- function(x, n = 2){stats::filter(x, rep(1 / n, n), sides = 2)}
plot_lag_with_6dayL <- as.list(rep(NA,length=unique(combo$Site)))

for(s in 1:length(unique(combo$Site))){
  dd <- combo[combo$Site==unique(combo$Site)[s],]
  
  t <- ma(x=dd$temp_sst,n=6)
  b <- ma(x=dd$temp_bottom,n=6)
  d <- as.data.frame(cbind(dd$Date,t,b))
  sst <- d$t[1:(length(d$b)-best_lags[best_lags$site==dd$Site[1],]$dayslag)] #SST at t
  bottom <- d$b[(best_lags[best_lags$site==dd$Site[1],]$dayslag+1):length(d$b)] #bottom at t+n
  day <- seq(from=1,to=length(sst),by=1)
  dplot <- as.data.frame(cbind(day,sst,bottom))
  dplot <- dplot[complete.cases(dplot),]
   
  plot_lag_with_6dayL[[s]]<- ggplot(data=dplot) + 
    geom_line(aes(x=day,y=sst,color="sst")) +
    geom_line(aes(x=day,y=bottom,color="bottom")) +
    xlab("") + ylab("") + 
    theme(legend.position="none") +
    ggtitle(paste(dd$Site[1]))   
}
rm(dd,t,b,d,sst,bottom,day,dplot,s)
```  
 
### Fig 4. Six day moving average window & adjusted time lag for each site

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=14, fig.width=7}
do.call(grid.arrange,c(plot_lag_with_6dayL,ncol=2,left="Temperature",bottom="Day Index"))
``` 

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# plot spectrum of raw SST, bottom
# plot spectrum of moving average SST, bottom with 6 day window
# setting 'span' - a vector of odd integers to specify the smoothers
ma <- function(x, n = 2){stats::filter(x, rep(1 / n, n), sides = 2)}
span.multiplier = 1
specplots <- as.list(rep(NA,length=length(unique(combo$Site))))
names(specplots) <- unique(combo$Site)

for(s in 1:length(specplots)){
  # subset to a site
  dd <- combo[combo$Site==unique(combo$Site)[s],]
    # square root time series length
  tmp <- ceiling(sqrt(length(1:length(dd$temp_bottom)))) #sq root of timeseries lgth, rounded
  if (tmp %% 2 == 0) {m <- tmp+1} else {m <- tmp} #make it odd, if the square root is even
  m = m * span.multiplier
  
  # sst raw spec
  sp1 <- spec.pgram(dd$temp_sst,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p1 <- ggplot() + geom_line(aes(x=sp1$freq,y=log10(sp1$spec))) +
    ggtitle("raw SST ") + xlab("freq")
  
  # bottom raw spec
  sp2 <- spec.pgram(dd$temp_bottom,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p2 <- ggplot() + geom_line(aes(x=sp2$freq,y=log10(sp2$spec))) +
    ggtitle("raw bottom ") + xlab("freq")
  
  # sst smooth spec
  sstS <- ma(dd$temp_sst,n=6)
  sstS <- sstS[!is.na(sstS)]
  sp3 <- spec.pgram(sstS,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p3 <- ggplot() + geom_line(aes(x=sp3$freq,y=log10(sp3$spec))) +
    ggtitle("smooth SST") + xlab("freq")
  #ggtitle(paste(dd$Site[1]," - smooth SST ",sep="")) + xlab("freq")
  
  # bottom smooth spec
  bottomS <- ma(dd$temp_bottom,n=6)
  bottomS <- bottomS[!is.na(bottomS)]
  sp4 <- spec.pgram(bottomS,spans=NULL,taper=0.1,plot = FALSE,demean = TRUE)
  p4 <- ggplot() + geom_line(aes(x=sp4$freq,y=log10(sp4$spec))) +
    ggtitle("smooth bottom ") + xlab("freq")
  
  siteplots <- list(p1,p2,p3,p4)
  
  specplots[[s]] <- siteplots
  
}

```

### Spectrums

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[1]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[1])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[2]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[2])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[3]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[3])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[4]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[4])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[5]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[5])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[6]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[6])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[7]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[7])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[8]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[8])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[9]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[9])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[10]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[10])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[11]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[11])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[12]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[12])))
```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
do.call(grid.arrange,c(specplots[[13]],ncol=2,left="",bottom="",top=paste(unique(combo$Site)[13])))
```